Notes on filtration of logics containing {\bf K5}

### 2021-06-04 filtration-stuff/NewText.tex
proof for K5 * K5

@@

    $\Delta_1 = V \cup \operatorname{Sub}(\{ \Diamond \Box p_{\psi} \: | \: \Box_1 \psi \in \Gamma \})$

    $\Delta_2 = V \cup \operatorname{Sub}(\{ \Diamond \Box p_{\psi} \: | \: \Box_2 \psi \in \Gamma \})$
  \end{center}
  Let us define equivalence relations $\sim_1$ and $\sim_2$ induced by $\Delta_1$ and $\Delta_2$ respectively.

  Consider $\mathcal{M}_i = \langle W, R_i, \vartheta' \rangle$, a reduct of $\mathcal{M}'$, we have:
  \begin{enumerate}
    \item $\mathcal{M}_i, x \models \Box p_{\psi}$ iff $\mathcal{M}, x \models \Box_i \psi$
    \item $\mathcal{M}_i, x \models \Diamond \Box p_{\psi}$ iff $\mathcal{M}, x \models \Diamond_i \Box_i \psi$
  \end{enumerate}
  So $\sim_i = \sim_{\Gamma'}$ by the construction. Let us put $\widehat{W} = W / \sim_{\Gamma'}$.

  \begin{claim}
    Let $\widehat{R}_i = (R^{min}_{\Delta_i})^{E}$ and
    $\widehat{\vartheta(p)} = \{ [x]_{\sim_i} \: | \: \mathcal{M}_i, x \models p \}$ for $p \in \Delta_1$, define $\widehat{\mathcal{M}}_i = \widehat{W}, \widehat{R}_i, \widehat{\vartheta}$. Then $\widehat{\mathcal{M}}_i \models {\bf K}5$ and $\widehat{\mathcal{M}}_i$ is a filtration of $\mathcal{M}_i$ through $\Delta_i$.
  \end{claim}
  \begin{proof}
    Let us show that $\widehat{R}_i \subseteq R^{max}_{\Delta_i}$ by induction.
    Let $\hat{x} \widehat{R}_i \hat{y}$ such that $\mathcal{M}_i, x \models \Box p_{\psi}$.
    We need $\mathcal{M}_i, y \models p_{\psi}$.
    Recall that $\widehat{R}_i$ is defined inductively as:
    \begin{enumerate}
      \item $R_{\Delta_i}^{0} = R_{\Delta_i}^{min}$
      \item $R_{\Delta_i}^{n + 1} = R_{\Delta_i}^{n} \cup ({R_{\Delta_i}^{n}}^{-1} \circ R_{\Delta_i}^{n})$
      \item $(R^{min}_{\Delta_i})^{E} = \bigcup_{k < \omega} R_{\Delta_i}^{k}$
    \end{enumerate}

    \begin{enumerate}
      \item $n = 0$. This is obvious.
      \item $m = n + 1$. Suppose the statement holds for $R_{\Delta_i}^{n}$, we also have $(\hat{x}, \hat{y}) \in R_{\Delta_i}^{n} \cup ({R_{\Delta_i}^{n}}^{-1} \circ R_{\Delta_i}^{n})$.
      If $(\hat{x}, \hat{y}) \in R_{\Delta_i}^{n}$, then we just apply IH.

      $(\hat{x}, \hat{y}) \in {R_{\Delta_i}^{n}}^{-1} \circ R_{\Delta_i}^{n} \in R_{\Delta_i}^{n}$, then there exists $\hat{z}$ such that $(\hat{z}, \hat{x}) \in R_{\Delta_i}^{n}$ and $(\hat{z}, \hat{y}) \in R_{\Delta_i}^{n}$.

      We rewrite $(\hat{z}, \hat{x}) \in R_{\Delta_i}^{n}$ as the following sausage (for some $\widehat{u_1},  \widehat{u_2}, \dots, \widehat{u_{n - 1}}, \widehat{u_{n}}$):

\xymatrix{
&&& \hat{z} & \ar[l]_{R_{\Delta_i}^{min}} \ar[r]^{R'} \widehat{u_1} & \ar[r]^{R'} \widehat{u_2} \ar[r]^{R'} & \dots & \ar[r]^{R'} \widehat{u_{n - 1}} \ar[r]^{R'} & \widehat{u_{n}} \ar[r]^{R_{\Delta_i}^{min}} & \hat{x}
}
\vspace{\baselineskip}

where $R'$ is either $R_{\Delta_i}^{min}$ or its converse. We clearly have $\mathcal{M}, z \models \Box p_{\psi}$, since $\mathcal{M}, x \models \Box p_{\psi}$ implies $\mathcal{M}, u_{n} \models \Diamond \Box p_{\psi}$. After that we apply the following property of ${\bf K}5$-models:
\begin{center}
  For each $a, b \in M_i$ such that $a R_i b$ we have $\mathcal{M}_i, a \models \Diamond \Box p_{\psi}$ iff $\mathcal{M}_i, b \models \Diamond \Box p_{\psi}$
\end{center}
Then $\mathcal{M}_i, y \models p_{\psi}$ by IH.
    \end{enumerate}
  \end{proof}

@@

  Let $\hat{x} \widehat{R_{\Gamma'}}_i \hat{y}$ and $\mathcal{M}, x \models \Box_i \psi$ for $\Box_i \psi \in \Gamma$. Then $\mathcal{M}_i, x \models \Box p_{\psi}$, so $\widehat{\mathcal{M}_i}, \hat{x} \models \Box p_{\psi}$.
  By the claim above, $\widehat{\mathcal{M}_i}$ is a filtration of $\mathcal{M}_i$ through $\Delta_i$, so $\mathcal{M}_i, y \models p_{\psi}$. Then $\mathcal{M}, y \models \psi$.

  \item Likewise
  \item
  \end{enumerate}
\end{proof}

>>>>

\begin{lemma}\label{extension}
  Let $\Gamma$ be a finite and $\operatorname{Sub}$-closed set of formulas and $\mathcal{M} = \langle W, R, \vartheta \rangle$. Consider $\Gamma' = \Gamma \cup \{ \Diamond \Box \psi \: | \: \Box \psi \in \Gamma \}$. Let $\Delta$ be any finite and $\operatorname{Sub}$-closed extension of $\Gamma'$. Then a model $\widehat{M} = \langle W / \sim_{\Delta}, (R^{min}_{\Delta})^{E}, \widehat{\vartheta} \rangle$ is a filtration of $\mathcal{M}$ through $\Delta$.
\end{lemma}

\begin{proof}
  Recall that $(R^{min}_{\Delta})^{E}$ is defined inductively as:
  \begin{enumerate}
    \item $R_{\Delta}^{0} = R_{\Delta}^{min}$
    \item $R_{\Delta}^{n + 1} = R_{\Delta}^{n} \cup ({R_{\Delta}^{n}}^{-1} \circ R_{\Delta}^{n})$
    \item $(R^{min}_{\Delta})^{E} = \bigcup_{k < \omega} R_{\Delta}^{k}$
  \end{enumerate}

  If $R^{min}_{\Delta}$ is already a subrelation of $R^{max}$, so the base case is self-evident.

  Suppose the statement holds for $R_{\Delta}^{n}$, $(\hat{x},\hat{y}) \in R_{\Delta}^{n + 1}$, that is, there exists $\hat{z}$ such that $(\hat{z}, \hat{x}), (\hat{z}, \hat{y}) \in R_{\Delta}^{n}$.
  Let $\mathcal{M}, x \models \Box \psi$ (for $\Box \psi \in \Gamma$).

  We rewrite $(\hat{z}, \hat{x}) \in R_{\Delta}^{n}$ as the following sausage (for some $\widehat{u_1},  \widehat{u_2}, \dots, \widehat{u_{n - 1}}, \widehat{u_{n}}$):

  \xymatrix{
  &&& \hat{z} & \ar[l]_{R_{\Delta}^{min}} \ar[r]^{R'} \widehat{u_1} & \ar[r]^{R'} \widehat{u_2} \ar[r]^{R'} & \dots \ar[r]^{R'} & \widehat{u_{n - 1}} \ar[r]^{R'} & \widehat{u_{n}} \ar[r]^{R_{\Delta}^{min}} & \hat{x}
  }
  where $R'$ is either $R_{\Delta}^{min}$ or its converse. We have $\mathcal{M}, z \models \Box \psi$, since $\mathcal{M}, x \models \Box \psi$ implies $\mathcal{M}, u_{n} \models \Diamond \Box \psi$. After that we apply the following property of ${\bf K}5$-models:
  \begin{center}
    For each $a, b \in M_i$ such that $a R_i b$ we have $\mathcal{M}_i, a \models \Diamond \Box p_{\psi}$ iff $\mathcal{M}_i, b \models \Diamond \Box p_{\psi}$
  \end{center}
  Note that we always stay within $\Delta$ since $\Gamma' \subseteq \Delta$.
  Finally, we conclude $\mathcal{M}, x \models \psi$ from IH and $\mathcal{M}, z \models \Box \psi$.
\end{proof}

@@

    $\Delta = V \cup \operatorname{Sub}(\{ \Diamond \Box p_{\psi} \: | \: \Box_i \psi \in \Gamma, i n= 1,2 \})$
  \end{center}
  Let us define an equivalence relation $\sim_{\Delta}$ induced by $\Delta$.

  Consider $\mathcal{M}_i = \langle W, R_i, \vartheta' \rangle$, a reduct of $\mathcal{M}'$, we have:
  \begin{enumerate}
    \item $\mathcal{M}_i, x \models \Box p_{\psi}$ iff $\mathcal{M}, x \models \Box_i \psi$
    \item $\mathcal{M}_i, x \models \Diamond \Box p_{\psi}$ iff $\mathcal{M}, x \models \Diamond_i \Box_i \psi$
  \end{enumerate}
  So $\sim = \sim_{\Gamma'}$ by the construction. Let us put $\widehat{W} = W / \sim_{\Gamma'}$. Lemma~\ref{extension} implies the following claim:

  \begin{claim}
    Let $\widehat{R}_i = (R^{min}_{\Delta})^{E}$ and
    $\widehat{\vartheta(p)} = \{ [x]_{\sim_i} \: | \: \mathcal{M}_i, x \models p \}$ for $p \in \Delta_1$, define $\widehat{\mathcal{M}}_i = \widehat{W}, \widehat{R}_i, \widehat{\vartheta}$. Then $\widehat{\mathcal{M}}_i \models {\bf K}5$ and $\widehat{\mathcal{M}}_i$ is a filtration of $\mathcal{M}_i$ through $\Delta$.
  \end{claim}

@@

  Let $\hat{x} \widehat{R_{\Gamma'}}_i \hat{y}$ and $\mathcal{M}, x \models \Box_i \psi$ for $\Box_i \psi \in \Gamma$. Then $\mathcal{M}_i, x \models \Box p_{\psi}$, so $\widehat{\mathcal{M}_i}, \hat{x} \models \Box p_{\psi}$.
  By the claim above, $\widehat{\mathcal{M}_i}$ is a filtration of $\mathcal{M}_i$ through $\Delta$, so $\mathcal{M}_i, y \models p_{\psi}$. Then $\mathcal{M}, y \models \psi$.

  \item Likewise
  \end{enumerate}
\end{proof}
